Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-03-15 ribose-paper.RMD
last updates discussion

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated protein, compared to RT with placebo. Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and -recovery, or total training session volume. At one time point, RT with glucose led to less reduction in strength compared to placebo, and levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose. Further, there were no differences in daily macronutrient intake on pairwise consecutive days. This would suggest the effects of high vs. low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous in vitro studies [@mariappan11; @tanaka15] are not translatable to acute effects in human skeletal muscle in vivo.

@@

The results from the strength tests do indicate that the participants might have been fatigued during the intervention, as they exercised every day for 12 days straight, without breakfast other than the treatments supplied. In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing.  However, this was not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions. Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, seeing as glucose is the preferred energy source during strenuous activity [@mul15]. <!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response, thus less performance. This could also explain the differences between treatments in content of the stress responsive c-Myc [SOURCES], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc. If this was the case, it does seem inconclusive that strength did no differ between treatments at earlier or later time points. In addition, there were no differences in training volume per session between treatments, therefore, the mechanical loading should have been equal. Another potential explanation to this observation might be …. (bifasisk restitusjon?)

Despite a thoroughly standardised design, the lack of participants do present a weakness of the present study. Initially, the minimum limit for statistical power was reached however, three participants dropped out during the intervention. Coupled with the fact that no previous study could guide our power calculation, this does introduce a degree of uncertainty regarding our results. Moreover, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective. Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them. If further investigations are to be made on this topic, these analyses should include a greater sample size, with a negative control. 
In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following two weeks of heavy resistance training in moderately trained young male and female adults. Neither does it acutely enhance muscular performance or recovery withing 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque.

>>>>

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated protein, compared to RT with placebo. Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and -recovery, or total training session volume. At one time point, RT with glucose led to less reduction in strength compared to placebo, and levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose. Further, there were no differences in daily macronutrient intake on pairwise consecutive days. This would suggest the effects of high vs. low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous in vitro studies [@mariappan11; @tanaka15] are not translatable to acute effects in human skeletal muscle in vivo. Notably, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective. Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.

@@

The results from the strength tests do indicate that the participants might have been fatigued during the intervention, as they exercised every day for 12 days straight, with strength testing in-between. In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing compared to placebo.  However, this was not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions. Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, seeing as glucose is the preferred energy source during strenuous activity [@mul15]. <!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance in strength tests. This could also explain the differences between treatments in content of the stress responsive c-Myc [SOURCES], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc. If this was the case, it does seem inconclusive that strength did not differ between treatments at earlier or later time points. In addition, there were no differences in training volume per session between treatments, therefore, the mechanical loading should have been equal. Another potential explanation to this observation might be …. (bifasisk restitusjon?). Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design. Initially, we met our minimum limit of 16 participants however, three dropped out during the intervention, leaving us slightly underpowered according to the a priori calculation.

In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following two weeks of heavy resistance training in moderately trained young male and female adults. Neither does it acutely enhance muscular performance or recovery withing 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque. The observations in total RNA, rRNA and rDNA transcription initiation associated proteins confirm the key role of increased UBF content in ribosome biogenesis in moderately trained human skeletal muscle following resistance training, previously observed in RT-untrained participants, rodents or *in vitro* [SOURCES?]. If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher resolution results.
